---
title: "R Notebook"
output: html_notebook
---

#Peat depth predictions

This file mainly does a cross validation analysis of the peat depth data
with the aim of comparing predictive accuracy for a number of models

------------------------------------------------------------------------------
File authors; Duncan Lee (University of Glasgow); Dylan Young (University of
Leeds)

Data from Dartmoor provided by Lauren Parry. A version of this file was used in
Spatial models with covariates improve estimates of peat depth in blanket
peatlands 
Young, Dylan M. & Parry, Lauren & Lee, Duncan & Ray, Surajit. (2018). 
Spatial models with covariates improve estimates of peat depth in blanket 
peatlands. PLOS ONE. 13. e0202691. 10.1371/journal.pone.0202691. 

Please give credit to the originators and make clear any modifications you make
(including the date the modifications were made).

------------------------------------------------------------------------------

### Load libraries required.
```{r}
library(geoR)
library(tidyverse)
```

------------------------------------------------------------------------------

## 1. Prepare data

Set the column numbers for covariates

5 - elevation
6 - slope
7 - Aspect
8 - Aspect category
9 - Vegetation
10 - Soil

```{r}
covars <- c(5, 6)

# Read in the data
#dat <- read.csv(file = "peat_depth_observations_dartmoor.csv")
dat <- read.csv(file = "../young_2018/dataset_604.csv")
dat.bak <- dat

# Choose which data to use in the predictions; ST = stratified sampling, GR =
# gridded sampling. If no choice is made then all data is used. Edit as necessary
# Stratified sampling
dat.choice <- 'ST'
# 250 m grid sampling
dat.choice <- 'GR'

if (!exists('dat.choice')) {
  dat <- dat
} else if (dat.choice == 'ST') {
  dat <- dat[dat$Source %in% 'ST', ]
  dat$Source <- factor(dat$Source)
} else if (dat.choice == 'GR') {
  dat <- dat[dat$Source %in% 'GR', ]
  dat$Source <- factor(dat$Source)
}

# Normalise the data
#  In exploratory analyses sqrt gave more normal looking residuals than log.
dat$sqrtdepth <- sqrt(dat$depth)
head(dat)
n <- nrow(dat)

```


------------------------------------------------------------------------------

## 2. Exploratory analysis - uncomment as necessary

### boxplots and pairs plots of the data

```{r}

 # pairs(dat[ ,c(19, 1,2,5,6,7,11)], pch=19)
 # pairs(dat[ ,c(19, 12:18)], pch=19)
 # boxplot(dat$sqrtdepth~dat$Source)
 # boxplot(dat$sqrtdepth~dat$Aspect_cat)
 # boxplot(dat$sqrtdepth~dat$Vegetation)
 # boxplot(dat$sqrtdepth~dat$Soil)

```


NOTE: aspect is circular need to account for this - sine or cosine or arc sine
transformation of it.

### Exploratory linear model analysis

```{r}

mod.lm.test <- lm(sqrtdepth~elevation + Slope_5m, data=dat)
summary(mod.lm.test)

resid <- residuals(mod.lm.test)
plot(resid)
qqnorm(resid)
qqline(resid)
dat$resid <- resid
```

Sqrt scale looks ok

------------------------------------------------------------------------------

## 3. Fit a geostatistical model to the data to look for evidence of spatial autocorrelation after accounting for covariates

 Create a geoR data object - the package that does the spatial modelling
 this requires specifying the response, coordinates and covariates.
 The covariates are based on the model above, if you want different ones you
 will need to change them in covars

```{r}
sp_depth <- as.geodata(obj=dat, coords.col=1:2, data.col=11, covar.col = covars)

# Jitter the 2 duplicated data locations which appear to be identical
names(sp_depth)
#coords, data, covariate (source, elevation)
set.seed(2490)
sp_depth <- jitterDupCoords(sp_depth, max=0.01)
```

### Plot the spatial object
In the top left plot the data have been split into 4 colours based on
quartiles (red is highest)

```{r}
plot(sp_depth)
```


### Fit a geostatistical model to all the data

```{r}
model.geo <- likfit(geodata=sp_depth, trend=~elevation + Slope_5m,
                    ini.cov.pars=c(15, 0.05), fix.nugget=FALSE,
                    cov.model="exponential")
summary(model.geo)
```

```{r}
plot(variog(sp_depth), main = "why does line and variog not line up?")
lines(model.geo)
```


The key covariance parameters here are:

- partial sill - amount of spatially correlated variation
- nugget - amount of non-spatial random variation
- asymptotic range - distance in degrees two point have to be apart for them to be uncorrelated.
- partial sill > nugget here so strong spatial variation after accounting for
 covariates.

------------------------------------------------------------------------------

## 4.  Conduct the cross validation exercise

 Split the data into 10 roughly equally sized groups, and use 9 groups to fit
 the model and predict sqrt depth for the tenth group. Repeat this for each of
 the 10 groups in turn.
 This may depend on how the data are split into 10 groups, so repeat this 10
 times with different groupings to average over this variation.

 WARNING - this will take a while (hours) to run for large datasets.

 The code below runs two models:
 
 - a simple linear model with given covariates.
 - a spatial correlation model with the same covariates.
 - common spatial covariance models are exponential, spherical and matern.
 
 This is a non-Bayesian analysis.

#### Create a matrix to store the results
 Store for each model the prediction and a 95% prediction interval
 Fit two models, a linear model with no spatial correlation (LM) and a full
 spatial correlation model with the same mean function (SM).
 LPI and UPI stand for lower / upper prediction interval.

```{r}
## Array - rows, columns, number of arrays
results <- array(NA, c(n, 6,10))
colnames(results) <- c("pred_LM", "LPI_LM", "UPI_LM", 
                       "pred_SM", "LPI_SM", "UPI_SM")
```

 One slight issue with geoR is when it predicts with factor (categorical)
 covariates. If the prediction set does not include all combinations of any
 factor there will be an error.
 Therefore set up a set of dummy prediction points that contain all
 combinations of the factors.
 Predictions at these dummy points will not be used.
 The covariates included are soil, source and vegetation.
 The locations of these dummy locations need to be jittered to avoid numerical
 issues.
 
```{r}
#Which are the columns with factors?
dat.fact <- dat[ ,sapply(dat, is.factor)]
#We don't need the Aspect_cat
dat.fact <- dat.fact[ ,-2]
#Keep only the unique rows
dat.fact <- dat.fact[!duplicated(dat.fact), ]
dummy <- dat[c(1, 2, 18, 106), ]
dummy <- dat[row.names(dat.fact), ]
dummy$longitude <- dummy$longitude + rnorm(dim(dummy)[1], sd = 0.01)
dummy$latitude <- dummy$latitude + rnorm(dim(dummy)[1], sd = 0.01)
n.dummy <- nrow(dummy)
```

#  
#  
#### The big loop:  
```{r}

# Loop over the 10 replications of splitting the data into 10 equally size groups.

nreps <- 10

for(r in 1:nreps) {
  print(r)
# Split the sites into 10 random groups of equal size.
 
  # Create random orderings until each of the 10 groups has at least one Winter Hill soil type
test=1
    while(test==1)
    {
    data.order <- sample(1:n)
    split_m_total <- length(data.order)
    split_m_G2 <- as.integer(split_m_total / 10)
    split_m_size <- length(data.order) %% 10
    # Therefore split_m_size chunks will have size split_m_G2 + 1 and the
    # remaining will have size split_size_G2
    split_m_G1 <- split_m_G2 + 1

    # For all data
    # To allow for changes in number of rows in dat
    split.matrix <-
      data.frame(sites=data.order,
        group = c(kronecker(1:split_m_size, rep(1,split_m_G1)),
                kronecker((split_m_size + 1):10, rep(1,split_m_G2))))
     # here column 1 is the row numbers and column 2 is the group number
     # Check each group has 2 Soil types
    types <- rep(NA,10)
        for (j in 1:10)
        {
        dat.temp <- dat[split.matrix[split.matrix$group == j,1] , ]
        types[j] <- sum(table(dat.temp$Soil) == nrow(dat.temp))
        }
    test <- max(types)
    }

# Undertake the 10-fold cross validation
   for (i in 1:10) {
     # Set up the fitting and the prediction data sets
    dat.fit <- dat[split.matrix[split.matrix$group!=i,1] , ]
    dat.pred <- dat[split.matrix[split.matrix$group==i,1] , ]
    dat.pred <- rbind(dat.pred, dummy)

     # Fit the linear model
    model.lm <- lm(formula=sqrtdepth~elevation + Slope_5m, data=dat.fit)
    model.lm.predictions <- 
      predict(object=model.lm, newdata=dat.pred, interval="prediction")

     # Fit the spatial model
       # Set up the spatial data object for the fitting data set
       # Ensure you add in the correct covariates here that you wish to use
    test.sp <-
      as.geodata(obj=dat.fit, coords.col=1:2, data.col=11, covar.col= covars)
     # Jitter anyduplicated data locations which appear to be identical
    test.sp <- jitterDupCoords(test.sp, max=0.01)


     # Fit the models
    model.sm <-
      likfit(geodata=test.sp, trend=~elevation + Slope_5m,
             ini.cov.pars=c(15, 0.05), fix.nugget=FALSE,
             cov.model="exponential", messages = F)
     # There are alternative spatial correlation models  so you could use these instead of exponential (e.g. spherical)
    print(r) # Monitor progress; each of the 10 r will be run 10(i) times.
    print(i)

     # Do the predictions
    control.sm <-
      krige.control(type.krige="OK", trend.d=~dat.fit$elevation +
                    dat.fit$Slope_5m, trend.l=~dat.pred$elevation +
                    dat.pred$Slope_5m, obj.model=model.sm)

    kriging.sm <-
      krige.conv(geodata=test.sp, locations=dat.pred[ ,1:2], krige=control.sm)

    model.sm.predictions <-
      cbind(kriging.sm$predict,
            kriging.sm$predict - 1.96 * sqrt(kriging.sm$krige.var),
            kriging.sm$predict+ 1.96 * sqrt(kriging.sm$krige.var))

     # Save the results but removing the last n.dummy values added to make the
     # kriging function work
    n.preds <- nrow(dat.pred) - n.dummy

    for(j in 1:n.preds)
    {
         # Choose the row (site) the  prediction corresponds to.
        which.row <- which(rownames(dat.pred)[j]==rownames(dat))

         # Save the results
        results[which.row, 1:3, r] <- model.lm.predictions[j, ]
        results[which.row, 4:6, r] <- model.sm.predictions[j, ]
    }
  }
}

```


#### Write predictions to dataframe

```{r}
#combine original data and predictions
results.df <- cbind(id = seq(1:nrow(dat)), dat, data.frame(results))
results.df
#make tidy data
results.ty <- gather(results.df, key = key, value = value, pred_LM.1:UPI_SM.10) %>% 
  separate(col = key, into = c("key", "modeltype", "run")) %>% 
  spread(key = key, value = value)
results.ty

```


------------------------------------------------------------------------------

## 5. Summarise the results  

### Compute metrics of interest  

#### Bias  

- measures whether the predictions are too large or too small on average
- want to be as close to zero as possible but due to random variation will not be exactly zero.

```{r}
# Linear model
mean(results[,1, ] - matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                            byrow=FALSE))
# Spatial model
mean(results[,4, ] - matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                            byrow=FALSE))
```


#### RMSE  

- measures average difference between the true and predicted values ignoring sign. 
- want to be as small as possible  

```{r}
# Linear model
sqrt(mean((results[,1, ] - matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                                  byrow=FALSE))^2)) 
# Spatial model
sqrt(mean((results[,4, ] - matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                                  byrow=FALSE))^2))

```
####  
#### Coverage 

- measures the probability that the 95% prediction intervals contain the true value 
- want to be 0.95


```{r}
# Linear model
mean(matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10, byrow=FALSE) >
     results[ ,2, ] & matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                             byrow=FALSE) < results[ ,3, ])
# Spatial model
mean(matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10, byrow=FALSE) >
     results[ ,5, ] & matrix(rep(dat$sqrtdepth,10), nrow=n, ncol=10,
                             byrow=FALSE) < results[ ,6, ])
```


#### Interval width   

- the width of the 95% prediction intervals
- want to be as small as possible provided that the coverage above is around 0.95. If the coverage is much lower than 0.95 then this is meaningless.

```{r}
# Linear model
mean(results[ ,3, ] - results[ ,2, ])
# Spatial model
mean(results[ ,6, ] - results[ ,5, ])
```

### visualise predictions

```{r}
graphdata <- results.ty %>% #filter(modeltype == "LM") %>% 
  group_by(modeltype, id) %>% 
  summarise(depth = mean(depth), pred = mean(pred)) #calculate mean result for each point

graphmetrics <- graphdata %>% 
  group_by(modeltype) %>% 
  summarise(RMSE = sqrt(mean(depth - (pred^2))^2),
            cc = cor(depth, (pred^2)))

ggplot(graphdata, aes(x = depth, y = pred^2)) +
  geom_point(aes(colour = modeltype))+
  geom_abline(slope = 1) +
  coord_equal() +
  geom_text(data = graphmetrics, aes(x = 50, y = c(200, 250), colour = modeltype, 
                                     label = paste("RMSE =", round(RMSE, 2), "cm \nCC =", round(cc, 2))))

```


```{r}
class(model.geo)
summary(model.geo)

```






------------------------------------------------------------------------------  

## 6. Do some plots

#### Plot the semi-variogram to test for the presence of spatial autocorrelation

```{r}
residuals <- residuals(model.lm)
dat.fit$residuals <- residuals
resid.sp <- as.geodata(obj=dat.fit, coords.col=1:2, data.col=11,
                       covar.col= covars)
resid.sp <- jitterDupCoords(resid.sp, max=0.01)
vari1 <- variog(resid.sp)
vari1.mc <- variog.mc.env(resid.sp, obj.variog=vari1)

plot(vari1, envelope.obj = vari1.mc, xlab="Distance (°)",
     ylab="Estimated semi-variogram", las = 1, cex = 1.3, cex.axis = 1.2,
     xaxs = 'i', yaxs = 'i', xlim = c(0,0.27), ylim = c(0,34), yaxt = 'n',
     cex.lab = 1.2)
axis(2, at = seq(0, 30, 10), labels = T, cex.axis = 1.2, las = 1)

```



## Try and use this model to predict against my own predictors

```{r}
library(geoR)
library(tidyverse)
library(sf)
library(sp)
library(gstat)
library(raster)
```

####Import predictors
```{r}
  #Import observations----
data.shp <- rgdal::readOGR("../data/peat_depths_dartmoor/peat_depths_dartmoor.shp")
head(coordinates(data.shp))
proj4string(data.shp)

data <- data.frame(data.shp)
#coords <- coordinates(data.shp) #eastnorth
#proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"
coords <- coordinates(sp::spTransform(data.shp, CRS("+proj=longlat +datum=WGS84"))) #lonlat
proj4 <- "+proj=longlat +datum=WGS84" #proj4 lonlat

data.sp <- sp::SpatialPointsDataFrame(data = data, coords = coords, proj4string = CRS(proj4))

  #import predictors from cropped raster stack----
elev <- raster("../data/predictor_stack.tif", band = 1)
slope <- raster("../data/predictor_stack.tif", band = 3)
  #create raster stack
predictors <- stack(elev, slope)
names(predictors) <- c("elevation", "Slope_5m")
#plot(predictors)
  #convert to spatial points dataframe
predictors.sp <- rasterToPoints(x = predictors, spatial = TRUE)

#select a well sampled area (of the NE data) to predict within----
# plot(input.data.sp)
crop.extent <- extent(c(xmin = 25575, xmax = 260000, ymin = 75000, ymax = 80000))
# plot(crop(input.data.sp, crop.extent))
predictors.cropped.sp <- crop(predictors.sp, crop.extent)
predictors.cropped.sp <-spTransform(predictors.cropped.sp, CRSobj = proj4string(data.sp)) #transform to lonlat if necessary
predictors.cropped.map <- data.frame(predictors.cropped.sp@data, predictors.cropped.sp@coords)

#crop data

```




```{r}
summary(predictors.sp)
names(predictors.sp)
spplot(predictors.sp["elevation"])
sp.points(data.sp["PEAT_DEPTH"])
spplot(predictors.cropped.sp["Slope_5m"])
```



```{r}
# # Fit the models
# model.sm <-
#   likfit(geodata=test.sp, trend=~elevation + Slope_5m,
#          ini.cov.pars=c(15, 0.05), fix.nugget=FALSE,
#          cov.model="exponential", messages = F)
# # There are alternative spatial correlation models  so you could use these instead of exponential (e.g. spherical)
# # Do the predictions
#    control.sm <-
#      krige.control(type.krige="OK", 
#                    trend.d=~dat.fit$elevation + dat.fit$Slope_5m, 
#                    trend.l=~dat.pred$elevation + dat.pred$Slope_5m, obj.model=model.sm)
# 
#    kriging.pred <-
#      krige.conv(geodata=test.sp, locations=dat.pred[ ,1:2], krige=control.sm)

# Fit the model
  model.sm <-
      likfit(geodata=data.sp, trend=~elevation + Slope_5m,
             ini.cov.pars=c(15, 0.05), fix.nugget=FALSE,
             cov.model="exponential", messages = F)
     # There are alternative spatial correlation models  so you could use these instead of exponential (e.g. spherical)
    
control.pred <-
  krige.control(type.krige="OK", 
                trend.d=~dat$elevation + dat$Slope_5m, 
                trend.l= ~ predictors.cropped.sp$elevation + predictors.cropped.sp$Slope_5m, 
                obj.model=model.geo)

kriging.pred <-
  krige.conv(geodata=sp_depth, 
             locations=predictors.cropped.sp@coords, 
             krige=control.pred)

prediction.sp <- predictors.cropped.sp
prediction.sp$pd_pred <-unlist( kriging.pred["predict"])

```


```{r}
spplot(prediction.sp["pd_pred"])
spplot(prediction.sp["elevation"])
spplot(prediction.sp["Slope_5m"])
```




## Test the model against our own data

To do this may need to check the following: 

- are the predictors the same - e.g. extract predictors for DY's data coordinates and compare

TO DO









