---
title: "Dartmoor code through"
output: html_notebook
---

Code through with an extract of data centred around dartmoor, based on the gstat Meuse example.  

##Prepare data

###Packages
```{r}
library(tidyverse)
library(sf)
library(sp)
library(gstat)
library(raster)
```

###Data

```{r}
data.shp <- sf::st_read("../data/peat_depths_dartmoor/peat_depths_dartmoor.shp")
summary(data.shp)
data <- data.frame(data.shp)
coords <- st_coordinates(data.shp)
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"

data.sp <- sp::SpatialPointsDataFrame(data = data, coords = coords, proj4string = CRS(proj4))

plot(data.sp)
```

```{r}
class(data.sp)
```


### Peat depths

```{r}
summary(data.sp$PEAT_DEPTH)
hist(data.sp$PEAT_DEPTH)
```

```{r}
names(data)
```


### Predictors

import cropped raster stack

```{r}
elev <- raster("../data/predictor_stack.tif", band = 1)
aspect <- raster("../data/predictor_stack.tif", band = 2)
slope <- raster("../data/predictor_stack.tif", band = 3)
# outflow <- raster("../data/predictor_stack.tif", band = 4)
# inflow <- raster("../data/predictor_stack.tif", band = 5)
# gdd <- raster("../data/predictor_stack.tif", band = 6)
# gsl <- raster("../data/predictor_stack.tif", band = 7)
# rain_ann <- raster("../data/predictor_stack.tif", band = 8)
# rain_daily <- raster("../data/predictor_stack.tif", band = 9)
# raindays_10mm <- raster("../data/predictor_stack.tif", band = 10)
# raindays_1mm <- raster("../data/predictor_stack.tif", band = 11)
# temp_mean <- raster("../data/predictor_stack.tif", band = 12)
# temp_min <- raster("../data/predictor_stack.tif", band = 13)
# temp_max <- raster("../data/predictor_stack.tif", band = 14)
```

create raster stack
```{r}
predictors <- stack(elev, aspect, slope)
names(predictors) <- c("ELEVATION", "ASPECT", "SLOPE")
plot(predictors)
```

```{r}
#convert to spatial points dataframe
predictors.sp <- rasterToPoints(x = predictors, spatial = TRUE)
crs(predictors.sp) <- crs(data.sp)
spplot(predictors.sp)
spplot(predictors.sp["ASPECT"])
```


###Extract predictor values at observation locations


```{r}
#create dataframe with peat depth as first col
input.data <- data.frame(peat_depth = data$PEAT_DEPTH)

#raster::extract environmental and topographic data for each peat depth measurement
input.data$elev <- raster::extract(
  elev, coordinates(data.sp)[,1:2])
input.data$aspect <- raster::extract(
  aspect, coordinates(data.sp)[,1:2])
input.data$slope <- raster::extract(
  slope, coordinates(data.sp)[,1:2])

#convert to spatial points data frame
coords <- st_coordinates(data.shp)
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"

input.data.sp <- sp::SpatialPointsDataFrame(data = input.data, coords = coords, proj4string = CRS(proj4))

spplot(input.data.sp)
```



##Interpolate

### Simple interpolation

Inverse distance weighting

```{r}
peatdepth.idw <- idw(peat_depth~1, input.data.sp, predictors.sp)
class(peatdepth.idw)
str(peatdepth.idw)
spplot(peatdepth.idw["var1.pred"], main = "peat depth inverse distance weighted interpolations")
```


### Variograms
```{r}
# logPD.vgm = variogram(log(peat)depth~1, input.data.sp)
# logPD.vgm
# logPD.fit = fit.variogram(logPD.vgm, model = vgm(1, "Sph", 900, 1))
# logPD.fit
# plot(logPD.vgm, logPD.fit)
```

NOT LOOKING GOOD, TRY WITHOUT LOG

```{r}
PD.vgm = variogram(peat_depth ~1, input.data.sp)
PD.vgm
plot(PD.vgm)
PD.fit.sph = fit.variogram(PD.vgm, model = vgm(500, "Sph", 2500, 1))
PD.fit.mat = fit.variogram(PD.vgm, model = vgm(500, "Mat", 2500, 1))
PD.fit.sph
plot(PD.vgm, PD.fit.sph)
plot(PD.vgm, PD.fit.mat)
```










Instead of the constant mean, denoted by \verb|~1|, we can specify a
mean function, e.g. using \verb|~sqrt(dist)| as a predictor variable:




```{r}
PD.regr.vgm = variogram(peat_depth ~ slope + elev, input.data.sp)
PD.regr.fit = fit.variogram(PD.regr.vgm, model = vgm(500, "Exp", 2500, 1))
PD.regr.fit
plot(PD.regr.vgm, PD.regr.fit)
```





Have a look at different co-variates

```{r}
PD.vgm.comp = data.frame(distance = variogram(peat_depth ~ 1, input.data.sp)$dist,
                         none = variogram(peat_depth ~ 1, input.data.sp)$gamma,
                         slope = variogram(peat_depth ~ slope, input.data.sp)$gamma,
                         elev = variogram(peat_depth ~ elev, input.data.sp)$gamma,
                         all = variogram(peat_depth ~ slope + elev, input.data.sp)$gamma)
                        
PD.vgm.comp = gather(PD.vgm.comp, key = covariate, value = semivariance, 2:5)

ggplot(PD.vgm.comp, aes(x = distance, y = semivariance))+
  geom_point(aes(shape = covariate))
```

##########################################
##from here on I'm using a different source for the input data - need to change that to the one that matches the predictor layers (i.e. input.data.sp not data.sp)

```{r}
PD.vgm.comp = rbind(data.frame(variogram(PEAT_DEPTH ~ 1, data.sp), covariate = "none"),
                    data.frame(variogram(PEAT_DEPTH ~ SLOPE, data.sp), covariate = "slope"),
                    data.frame(variogram(PEAT_DEPTH ~ ELEVATION, data.sp), covariate = "elev"),
                    data.frame(variogram(PEAT_DEPTH ~ consdryday, data.sp), covariate = "drydays"),
                    data.frame(variogram(PEAT_DEPTH ~ LCM2007, data.sp), covariate = "landcover"),
                    data.frame(variogram(PEAT_DEPTH ~ SLOPE + ELEVATION + consdryday + LCM2007, data.sp), 
                               covariate = "all.cont")
                    )
    
ggplot(PD.vgm.comp, aes(x = dist, y = gamma))+
  geom_point(aes(shape = covariate, colour = covariate))
```



```{r}
plot(variogram(PEAT_DEPTH ~ 1, data.sp, cloud = TRUE))
```



```{r}
head(data.sp)
table(data.sp$PEATTYPE)
table(data.sp$LCM2007)
class(data.sp$LCM2007)
```

###Kriging

First remove duplicate locations otherwise you get the following error: 
'Covariance matrix singular at location [276400,95250,0]: skipping...'

```{r}
zerodist(obj = data.sp, zero = 0, unique.ID = FALSE, memcmp = TRUE)
data.sp.dedup <- data.sp[-zerodist(data.sp)[,1],] 
dim(data.sp); dim(data.sp.dedup)
```



Ordinary kriging
```{r}
pd.kriged = krige(formula = PEAT_DEPTH~1, data.sp.dedup, predictors.sp, model = PD.fit.sph)
spplot(pd.kriged["var1.pred"])
spplot(pd.kriged["var1.var"])

summary(pd.kriged)
hist(pd.kriged$var1.pred)
hist(data$PEAT_DEPTH)
```

###Conditional simulation

```{r}
peatdepth.condsim = krige(PEAT_DEPTH~1, data.sp.dedup, predictors.sp, model = PD.fit.sph, nmax = 30, nsim = 4)
str(peatdepth.condsim)
spplot(peatdepth.condsim, main = "four conditional simulations")

```

  

###Cross Variography
Fitting a linear model of coregionalization.

```{r}
g = gstat(g = NULL, id = "peat-depth", formula = PEAT_DEPTH ~ 1, data = data.sp)
g = gstat(g, "elev", formula = ELEVATION ~ 1, data = data.sp)
v = variogram(g)
g = gstat(g, model = vgm(1, "Exp", 300, 1), fill.all = TRUE)
g.fit = fit.lmc(v, g)
g.fit
plot(v, g.fit)
vgm.map = variogram(g, cutoff = 1500, width = 100, map = TRUE)
plot(vgm.map, threshold = 5, col.regions = bpy.colors(), xlab = "", ylab = "")
```




