---
title: "Dartmoor code through"
output: html_notebook
---

Code through with an extract of data centred around dartmoor, based on the gstat Meuse example.  

##Prepare data

###Packages
```{r}
library(tidyverse)
library(sf)
library(sp)
library(gstat)
library(raster)
```

###Data

```{r}
data.shp <- sf::st_read("../data/peat_depths_dartmoor/peat_depths_dartmoor.shp")
summary(data.shp)
data <- data.frame(data.shp)
coords <- st_coordinates(data.shp)
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"

data.sp <- sp::SpatialPointsDataFrame(data = data, coords = coords, proj4string = CRS(proj4))

plot(data.sp)
```

```{r}
class(data.sp)
```


### Peat depths

```{r}
summary(data.sp$PEAT_DEPTH)
hist(data.sp$PEAT_DEPTH)
```

```{r}
names(data)
```


### Predictors

import cropped raster stack

```{r}
elev <- raster("../data/predictor_stack.tif", band = 1)
aspect <- raster("../data/predictor_stack.tif", band = 2)
slope <- raster("../data/predictor_stack.tif", band = 3)
```

create raster stack
```{r}
predictors <- stack(elev, aspect, slope)
names(predictors) <- c("elev", "aspect", "slope")
plot(predictors)
```

```{r}
#convert to spatial points dataframe
predictors.sp <- rasterToPoints(x = predictors, spatial = TRUE)
crs(predictors.sp) <- crs(data.sp)
spplot(predictors.sp)
spplot(predictors.sp["elev"])
```


###Extract predictor values at observation locations


```{r}
#create dataframe with peat depth as first col
input.data <- data.frame(peat_depth = data$PEAT_DEPTH)

#raster::extract environmental and topographic data for each peat depth measurement
input.data$elev <- raster::extract(
  elev, coordinates(data.sp)[,1:2])
input.data$aspect <- raster::extract(
  aspect, coordinates(data.sp)[,1:2])
input.data$slope <- raster::extract(
  slope, coordinates(data.sp)[,1:2])

#convert to spatial points data frame
coords <- st_coordinates(data.shp)
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"

input.data.sp <- sp::SpatialPointsDataFrame(data = input.data, coords = coords, proj4string = CRS(proj4))


spplot(input.data.sp)
```

```{r}
mapdata <- data.frame(input.data.sp@data, input.data.sp@coords)
head(mapdata)
```

```{r}
ggplot(mapdata, aes(x = X, y = Y)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  coord_equal()+
  labs(title = "Peat depth spatial distribution")
```

```{r}
ggplot(mapdata, aes(x = elev, y = slope)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  labs(title = "Peat depth: elevation by slope")
```

```{r}
ggplot(mapdata, aes(x = aspect, y = peat_depth)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  coord_polar() +
  labs(title = "Peat depth by aspect")

```

##Create prediction area

```{r}
#select a well sampled area to predict within
plot(input.data.sp)
crop.extent <- extent(c(xmin = 25575, xmax = 260000, ymin = 75000, ymax = 80000))
plot(crop(input.data.sp, crop.extent))
predictors.cropped.sp <- crop(predictors.sp, crop.extent)
predictors.cropped.map <- data.frame(predictors.cropped.sp@data, predictors.cropped.sp@coords)
```

```{r}
p.elev <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = elev)) + coord_equal() +
  scale_colour_distiller(palette = "Reds") +  
  labs(title = "Elevation", x = "", y = "")
p.asp <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = aspect)) + coord_equal() + 
  scale_colour_distiller(palette = "Blues") +  
  labs(title = "Aspect", x = "", y = "")
p.slo <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = slope))  + coord_equal()+ 
  scale_colour_distiller(palette = "Greens") +  
  labs(title = "Slope", x = "", y = "")
p.theme <- theme(legend.position = "bottom")

cowplot::plot_grid(
  p.elev + p.theme, 
  p.asp + p.theme,
  p.slo + p.theme,
  axis = "left", nrow = 1, labels = "auto"
)
```



##Interpolate

### Simple interpolation

Inverse distance weighting

```{r}
peatdepth.idw <- idw(peat_depth~1, input.data.sp, predictors.cropped.sp)
class(peatdepth.idw)
str(peatdepth.idw)
spplot(peatdepth.idw["var1.pred"], main = "peat depth inverse distance weighted interpolations")
```



### Variograms
```{r}
#log of peat depth plus one
logPD.vgm = variogram(log(peat_depth+1)~1, input.data.sp)
logPD.vgm
logPD.fit = fit.variogram(logPD.vgm, model = vgm(1, "Sph", 900, 1))
logPD.fit
plot(logPD.vgm, logPD.fit)
```



```{r}
PD.vgm = variogram(peat_depth ~1, input.data.sp)
PD.vgm
plot(PD.vgm)
PD.fit.sph = fit.variogram(PD.vgm, model = vgm(500, "Sph", 2500, 1))
PD.fit.mat = fit.variogram(PD.vgm, model = vgm(500, "Mat", 2500, 1))
PD.fit.sph
plot(PD.vgm, PD.fit.sph)
plot(PD.vgm, PD.fit.mat)
```



Instead of the constant mean, denoted by \verb|~1|, we can specify a
mean function, e.g. using \verb|~sqrt(dist)| as a predictor variable:

```{r}
PD.regr.vgm = variogram(peat_depth ~ slope + elev, input.data.sp)
PD.regr.fit = fit.variogram(PD.regr.vgm, model = vgm(500, "Exp", 2500, 1))
PD.regr.fit
plot(PD.regr.vgm, PD.regr.fit)
```





Have a look at different co-variates

```{r}
PD.vgm.comp = data.frame(distance = variogram(peat_depth ~ 1, input.data.sp)$dist,
                         none = variogram(peat_depth ~ 1, input.data.sp)$gamma,
                         slope = variogram(peat_depth ~ slope, input.data.sp)$gamma,
                         elev = variogram(peat_depth ~ elev, input.data.sp)$gamma,
                         all = variogram(peat_depth ~ slope + elev, input.data.sp)$gamma)
                        
PD.vgm.comp = gather(PD.vgm.comp, key = covariate, value = semivariance, 2:5)

ggplot(PD.vgm.comp, aes(x = distance, y = semivariance))+
  geom_point(aes(shape = covariate))
```

```{r}
plot(variogram(peat_depth ~ 1, input.data.sp, cloud = TRUE))
```


```{r}
head(input.data.sp)
```

###Kriging

First remove duplicate locations otherwise you get the following error: 
'Covariance matrix singular at location [276400,95250,0]: skipping...'

```{r}
zerodist(obj = input.data.sp, zero = 0, unique.ID = FALSE, memcmp = TRUE)
data.sp.dedup <- input.data.sp[-zerodist(data.sp)[,1],] 
dim(input.data.sp); dim(data.sp.dedup)
```



Ordinary kriging
```{r}
pd.kriged.crop = krige(formula = peat_depth ~ 1, data.sp.dedup, predictors.cropped.sp, model = logPD.fit)
saveRDS(pd.kriged.crop, "../data/pd.kriged.crop.logPd.rds")
pd.kriged.crop.logPd <- readRDS("../data/pd.kriged.crop.logPd.rds")

spplot(pd.kriged.crop.logPd["var1.pred"])
spplot(pd.kriged.crop.logPd["var1.var"])

```

regression kriging
```{r}

pd.kriged = krige(formula = peat_depth ~ 1, data.sp.dedup, predictors.cropped.sp, model = PD.regr.fit)
saveRDS(pd.kriged, "../data/pd.kriged.crop.regPd.rds")
pd.kriged.crop.regPd <- readRDS("../data/pd.kriged.crop.regPd.rds")

spplot(pd.kriged.crop.regPd["var1.pred"])
spplot(pd.kriged.crop.regPd["var1.var"])

summary(pd.kriged)
hist(pd.kriged$var1.pred)
hist(input.data$peat_depth)
```


###Conditional simulation

```{r}
pd.condsim = krige(peat_depth ~ 1, data.sp.dedup, predictors.cropped.sp, model = PD.fit.sph, nmax = 30, nsim = 4)
saveRDS(pd.condsim, "../data/pd.condsim.crop.sph.rds")
pd.condsim.crop.sph <- readRDS("../data/pd.condsim.crop.sph.rds")
spplot(pd.condsim.crop.sph, main = "four conditional simulations")

# pd.condsim.regr = krige(peat_depth ~ 1, data.sp.dedup, predictors.cropped.sp, model = PD.regr.fit, nmax = 30, nsim = 4)
# saveRDS(pd.condsim.regr, "../data/pd.condsim.regr.rds")
pd.condsim.regr <- readRDS("../data/pd.condsim.regr.rds")
spplot(pd.condsim.regr, main = "four conditional simulations")


```

  



