---
title: "Dartmoor workthrough with geoR"
author: "Christoph Kratz"
date: "24 August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Code through with an extract of data centred around dartmoor, using geoR vignette and approach from Young et al 2018.

###Packages
```{r}
library(geoR)
library(tidyverse)
library(sf)
library(sp)
library(gstat)
library(raster)
```

##Prepare data

```{r}
  #Import observations----
data.shp <- sf::st_read("../data/peat_depths_dartmoor/peat_depths_dartmoor.shp")
data <- data.frame(data.shp)
coords <- st_coordinates(data.shp)
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"
data.sp <- sp::SpatialPointsDataFrame(data = data, coords = coords, proj4string = CRS(proj4))

  #import predictors from cropped raster stack----
elev <- raster("../data/predictor_stack.tif", band = 1)
aspect <- raster("../data/predictor_stack.tif", band = 2)
slope <- raster("../data/predictor_stack.tif", band = 3)
  #create raster stack
predictors <- stack(elev, aspect, slope)
names(predictors) <- c("elev", "aspect", "slope")
#plot(predictors)
  #convert to spatial points dataframe
predictors.sp <- rasterToPoints(x = predictors, spatial = TRUE)
crs(predictors.sp) <- crs(data.sp)

#Extract predictor values at observation locations----
  #create dataframe with peat depth as first col
input.data <- data.frame(peat_depth = data$PEAT_DEPTH)
  #raster::extract environmental and topographic data for each peat depth measurement
input.data$elev <- raster::extract(
  elev, coordinates(data.sp)[,1:2])
input.data$aspect <- raster::extract(
  aspect, coordinates(data.sp)[,1:2])
input.data$slope <- raster::extract(
  slope, coordinates(data.sp)[,1:2])

#convert to spatial points data frame----
coords <- st_coordinates(data.shp)[,c("X", "Y")]
proj4 <- "+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +datum=OSGB36 +units=m +no_defs"

input.data.sp <- sp::SpatialPointsDataFrame(data = input.data, coords = coords, proj4string = CRS(proj4))
#spplot(input.data.sp)
```


####graph observations and environmental values

```{r}
mapdata <- data.frame(input.data.sp@data, input.data.sp@coords)
head(mapdata)
ggplot(mapdata, aes(x = X, y = Y)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  coord_equal()+
  labs(title = "Peat depth spatial distribution")

ggplot(mapdata, aes(x = elev, y = slope)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  labs(title = "Peat depth: elevation by slope")

ggplot(mapdata, aes(x = aspect, y = peat_depth)) +
  geom_point(aes(colour = peat_depth))+
  scale_color_distiller(palette = "Reds", direction = 1)+
  coord_polar() +
  labs(title = "Peat depth by aspect")
```

###Create prediction area

```{r}
#select a well sampled area to predict within----
# plot(input.data.sp)
crop.extent <- extent(c(xmin = 25575, xmax = 260000, ymin = 75000, ymax = 80000))
# plot(crop(input.data.sp, crop.extent))
input.data.cropped.sp <- crop(input.data.sp, crop.extent)
input.data.cropped.map <- data.frame(input.data.cropped.sp@data, 
                                     input.data.cropped.sp@coords) %>% 
  rename(x = X, y = Y)
predictors.cropped.sp <- crop(predictors.sp, crop.extent)
predictors.cropped.map <- data.frame(predictors.cropped.sp@data, predictors.cropped.sp@coords)
#graph it
p.theme <- theme(legend.position = "right")
p.elev <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = elev)) + coord_equal() +
  geom_text(data = input.data.cropped.map, label = "x", size = 0.8)+
  scale_colour_distiller(palette = "Reds") +  
  labs(#title = "Elevation", 
       x = "", y = ""); p.elev + p.theme
p.asp <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = aspect)) + coord_equal() + 
  geom_text(data = input.data.cropped.map, label = "x", size = 0.8)+
  scale_colour_distiller(palette = "Blues") +  
  labs(#title = "Aspect", 
       x = "", y = ""); p.asp + p.theme
p.slo <- ggplot(predictors.cropped.map, aes(x = x, y = y)) +
  geom_point(aes(colour = slope))  + coord_equal()+ 
  geom_text(data = input.data.cropped.map, label = "x", size = 0.8)+
  scale_colour_distiller(palette = "Greens") +  
  labs(#title = "Slope", 
       x = "", y = ""); p.slo + p.theme
# cowplot::plot_grid(p.elev + p.theme, p.asp + p.theme, p.slo + p.theme, axis = "left", nrow = 3, labels = "")
```

###Convert to inputs to geodata format for geoR

```{r}
input.data.to.use <- input.data.cropped.sp #spdf format
input.data.gd <- as.geodata(obj = input.data.to.use)
#remove duplicates 
dupes <- as.numeric(row.names(dup.coords(input.data.gd)))
cat("There are", length(dupes),"duplicate locations.  ")
if(length(dupes) > 0){
  input.data.gd <- as.geodata(obj = input.data.to.use[-dupes,])
  cat(length(dupes), "duplicates have been removed.")}
rm(input.data.to.use)
#in this case chosen to simply remove both instances of a duplicate record.  If duplicate locations are an important part of data then better to use jitterDupCoords
```


## 2. Descriptive plots

```{r}
plot(input.data.gd)
```

```{r}
max_dist <- 400
#I don't understand these plots
par(mfrow = c(2,2), mar=c(3,3,1,1), mgp=c(2,0.8,0))
points(input.data.gd, xlab = "Coord X", ylab = "Coord Y")
points(input.data.gd, xlab = "Coord X", ylab = "Coord Y", pt.divide = "rank.prop")
points(input.data.gd, xlab = "Coord X", ylab = "Coord Y", cex.max = 1.7, col = gray(seq(1, 0.1, l=max_dist)), pt.divide = "equal")
points(input.data.gd, pt.divide = "quintile", xlab = "Coord X", ylab = "Coord Y")
##dev.off()

```


##Variograms

###Empirical variograms
Empirical variograms are calculated using the function variog. There are options for the classical or modulus estimator. Results can be returned as variogram clouds, binned or smoothed variograms.


```{r}
max_dist <- 400
cloud1 <- variog(input.data.gd, option = "cloud")#, max.dist = max_dist)
cloud2 <- variog(input.data.gd, option = "cloud", estimator.type = "modulus")#, max.dist = max_dist)
bin1 <- variog(input.data.gd)#, uvec=seq(0, max_dist, l=11))
bin2  <- variog(input.data.gd, estimator.type= "modulus")#, uvec=seq(0, max_dist, l=11))

par(mfrow=c(2,2))
plot(cloud1, main = "cloud: classical estimator")
plot(cloud2, main = "cloud: modulus estimator")
plot(bin1, main = "binned: classical estimator")
plot(bin2, main = "binned: modulus estimator")
#par(par.ori)
```
	  

	    
```{r}
bin1 <- variog(input.data.gd, uvec = seq(0, max_dist,l=11), bin.cloud = T)
bin2 <- variog(input.data.gd, uvec = seq(0, max_dist,l=11), estimator.type = "modulus", bin.cloud = T)

##jpeg("input.data.gdplot/input.data.gdp04.jpeg", wid=450, hei=250)
par(mfrow=c(1,2),mar=c(3,3,2,2),mgp=c(2,.8,0))
plot(bin1, bin.cloud=T, main="classical estimator")
plot(bin2, bin.cloud=T, main="modulus estimator")
##dev.off()
par(par.ori)


```



```{r}
##
bin1 <- variog(input.data.gd, uvec = seq(0, max_dist, l=11))

##jpeg("input.data.gdplot/input.data.gdp05.jpeg", wid=250, hei=250)
par(mar=c(3,3,.2,.2),mgp=c(2,.8,0))
plot(bin1)
lines.variomodel(cov.model="exp", cov.pars=c(200,400), nugget = 200, max.dist = max_dist, lwd = 3)
smooth <- variog(input.data.gd, option = "smooth", max.dist = max_dist, n.points = 100, kernel = "normal", band = 0.2)
lines(smooth, type ="l", lty = 2)
legend(0.4, 0.3, c("empirical","exponential model","smooth"), lty=c(1,1,2), lwd=c(1,3,1), cex=0.7)
##dev.off()
par(par.ori)

```

